Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2019-12-06 xrb/paper.tex
Make signs for g consistent in HSE equation and parameter table.
Fix typo.
Add MarginPar asking to rename the Flame Acceleration section.

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) g
\end{equation}
and holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected at this boundary.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below), with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

@@

\begin{deluxetable}{lcc}
\tablecaption{\label{table:params} Initial model parameters.}
\tablehead{\colhead{parameter} & \colhead{cool} & \colhead{hot}}
\startdata
$T_\star$       & \multicolumn{2}{c}{$10^8$~K} \\
$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.4\times 10^9$~K \\
$T_\mathrm{lo}$ & \multicolumn{2}{c}{$8\times 10^6$~K} \\
$\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.36\times 10^6~\gcc$ & $1.20\times 10^6~\gcc$ \\
$\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
$H_\star$       & \multicolumn{2}{c}{2000~cm} \\
$\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
$X_\star(\isotm{Ni}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
\enddata
%
\tablenotetext{a}{This is not an input parameter, but instead is
  computed during integration.  We list it here for reference.}
%
\tablenotetext{b}{All other species are taken as 0.}
\end{deluxetable}

@@

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonable well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect, as we'll see in the 5/5 case below.

@@

To measure the propagation rate of the burning front, we first collapse our nuclear
energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
profile by averaging over the vertical coordinate. We then take the peak
$\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
We define the position of the flame front to be the location ahead of hottest part of
the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
This corresponds roughly to the leading edge of the burning region. Averaging
over the vertical coordinate helps to reduce sensitivity to localized fluid motions, as does
tracking the $0.1 \%$ contour rather than a local maximum. As we see in Figure
\ref{fig:flame_speed}, the flame settles into a state of steady propagation after an initial
transient period spanning $\sim 3$ ms. The position data here are well fitted by a linear
function of time, and the resulting slope gives the velocity of the flame front.

>>>>

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
    p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf y}
\end{equation}
and holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected at this boundary.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below), with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

@@

\begin{deluxetable}{lcc}
\tablecaption{\label{table:params} Initial model parameters.}
\tablehead{\colhead{parameter} & \colhead{cool} & \colhead{hot}}
\startdata
$T_\star$       & \multicolumn{2}{c}{$10^8$~K} \\
$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.4\times 10^9$~K \\
$T_\mathrm{lo}$ & \multicolumn{2}{c}{$8\times 10^6$~K} \\
$\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.36\times 10^6~\gcc$ & $1.20\times 10^6~\gcc$ \\
$\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
$H_\star$       & \multicolumn{2}{c}{2000~cm} \\
$\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
$X_\star(\isotm{Ni}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$\gb$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}} \hat{\bf y}$} \\
\enddata
%
\tablenotetext{a}{This is not an input parameter, but instead is
  computed during integration.  We list it here for reference.}
%
\tablenotetext{b}{All other species are taken as 0.}
\end{deluxetable}

@@

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonably well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect, as we'll see in the 5/5 case below.

@@

To measure the propagation \MarginPar{Can we rename this section "Flame Propagation"? The burning front in our simulations seems to move at a constant velocity, and we don't discuss acceleration.}
rate of the burning front, we first collapse our nuclear
energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
profile by averaging over the vertical coordinate. We then take the peak
$\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
We define the position of the flame front to be the location ahead of hottest part of
the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
This corresponds roughly to the leading edge of the burning region. Averaging
over the vertical coordinate helps to reduce sensitivity to localized fluid motions, as does
tracking the $0.1 \%$ contour rather than a local maximum. As we see in Figure
\ref{fig:flame_speed}, the flame settles into a state of steady propagation after an initial
transient period spanning $\sim 3$ ms. The position data here are well fitted by a linear
function of time, and the resulting slope gives the velocity of the flame front.
